---
title: "MA678 Final Project"
subtitle: "EDA, Feature Engineering and Model"
author: "Chenghao Meng"
date: "2020/12/1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,data.table,lubridate,magrittr,MASS,arm)
```

# 2 EDA and Feature Engineering
## 2.1 Explortary Data Analysis
```{r}
# Load the data without missing value
df <- fread(file = "C:/Users/CH.Meng/Desktop/MA678-Final project/hotel_imp2.csv")

# Date time transformation using lubridate
df$date_time <- as_datetime(df$date_time)
df$srch_ci <- as_datetime(df$srch_ci)
df$srch_co <- as_datetime(df$srch_co)

# Categorical transformation
df %<>% 
  dplyr::mutate_at(vars(
    c(site_name,posa_continent,user_location_country,
      user_location_region, user_location_city, user_id, 
      channel, srch_destination_id,srch_destination_type_id,
      hotel_continent,hotel_country,hotel_market,hotel_cluster)),~as.character(.))
```

### 2.1.1 Hotel-related

Which Hotel is most frequently booked?

```{r message=FALSE, warning=FALSE}
df %>% dplyr::select(is_booking,hotel_cluster) %>% 
  dplyr::filter(is_booking == 1) %>% 
  group_by(hotel_cluster) %>% 
  summarise(num=n()) %>% 
  arrange(desc(num)) %>% 
  
  # Top 10 booking hotel
  head(10) %>% ggplot() + 
  geom_bar(aes(x=reorder(hotel_cluster,num),y=num,fill=hotel_cluster),stat = "identity") + 
  scale_fill_brewer(palette = "PuOr") + 
  xlab("Hotel Cluster") + ylab("Frequency") + 
  ggtitle("Top 10 Most Frequently Booked Hotels",subtitle = "From Jan. 2013 to Dec. 2014") + 
  labs(fill="Hotel Cluster")+
  coord_flip()
```

Which is the most popular continent and country?

```{r message=FALSE, warning=FALSE}
# Bar plot: Hotel Continent
p_hotel_cont <- df %>% dplyr::select(is_booking,hotel_continent) %>% 
  group_by(hotel_continent) %>% summarise(num=n()) %>% 
  arrange(desc(num)) %>% 
  
  # Make the graph
  ggplot() + 
  geom_bar(aes(x=reorder(hotel_continent,num),y=num,fill=hotel_continent),stat = "identity") + 
  scale_fill_brewer(palette = "PuOr") + 
  xlab("Continent ID") + ylab("Number of Clicks") + 
  ggtitle("Clicks Frquency: Hotel Continent",subtitle = "From Jan. 2013 to Dec. 2014") + 
  labs(fill="Continent ID") + coord_flip()
p_hotel_cont
```

```{r message=FALSE, warning=FALSE}
# Bar plot: Top 10 Country
p_hotel_coun <- df %>% 
  dplyr::select(is_booking,hotel_country) %>% 
  group_by(hotel_country) %>% summarise(num=n()) %>% 
  arrange(desc(num)) %>% head(10) %>% 
  
  # Make the graph
  ggplot() + 
  geom_bar(aes(x=reorder(hotel_country,num),y=num,fill=hotel_country),stat = "identity") + 
  scale_fill_brewer(palette = "PuOr") + 
  xlab("Country ID") + ylab("Number of Clicks") + 
  ggtitle("Clicks Frquency: Top 10 Hotel Countries",subtitle = "From Jan. 2013 to Dec. 2014") + 
  labs(fill="Country ID") + coord_flip()
p_hotel_coun
```

### 2.1.2 User-related

#### Who booked hotel most frequently?

```{r message=FALSE, warning=FALSE}
df %>% dplyr::filter(is_booking==1) %>% 
  dplyr::select(user_id,is_booking) %>% 
  group_by(user_id) %>% summarise(num=n()) %>%
  arrange(desc(num)) %>% top_n(5,num) %>% 
  
  ggplot() + 
  geom_bar(aes(x=reorder(user_id,num),y=num),stat="identity",fill="royalblue",alpha=0.7)+
  xlab("User ID") + ylab("Number of Hotel Booking") + 
  ggtitle("Top 5 Users for Number of Hotel Booking",subtitle = "From Jan. 2013 to Dec. 2014") + 
  coord_flip()
```


#### Which country's user has the most hotel bookings?

```{r message=FALSE, warning=FALSE}
df %>% dplyr::select(is_booking,user_location_country) %>% 
  dplyr::filter(is_booking == 1) %>% 
  group_by(user_location_country) %>% 
  summarise(num=n()) %>% 
  arrange(desc(num)) %>% 
  
  head(10) %>% ggplot() + 
  geom_bar(aes(x=reorder(user_location_country,num),y=num,fill=user_location_country),stat = "identity") + 
  scale_fill_brewer(palette = "RdYlBu") + 
  xlab("Country ID the User is Located") + ylab("Frequency") + 
  ggtitle("Top 10 Countries for Number of Hotel Bookings per User",subtitle = "From Jan. 2013 to Dec. 2014") + 
  labs(fill="Country ID")+ coord_flip() 
```

#### Which hotel cluster the users from Top 5 countries for number of hotel bookings per user prefers to book?

```{r message=FALSE, warning=FALSE}
user_country_hotel <- df %>% 
  dplyr::select(is_booking,user_location_country,hotel_cluster) %>%
  dplyr::filter(is_booking==1,user_location_country==c("66","205","69","3","46")) %>% 
  group_by(user_location_country,hotel_cluster) %>% 
  summarise(num=n()) %>% 
  arrange(desc(num)) 
```

```{r}
# Top 5 hotel cluster the User from Top 5 countries for number of hotel bookings per user prefers to book
user_country66_hotel <- user_country_hotel %>% 
  filter(user_location_country=="66") %>% head(5)

user_country205_hotel <- user_country_hotel %>% 
  filter(user_location_country=="205") %>% head(5)

user_country69_hotel <- user_country_hotel %>% 
  filter(user_location_country=="69") %>% head(5)

user_country3_hotel <- user_country_hotel %>% 
  filter(user_location_country=="3") %>% head(5)

user_country46_hotel <- user_country_hotel %>% 
  filter(user_location_country=="46") %>% head(5)

user_country_hotel2 <- rbind(user_country66_hotel,user_country205_hotel,user_country69_hotel,user_country3_hotel,user_country46_hotel)
```


```{r message=FALSE, warning=FALSE}
ggplot(data=user_country_hotel2) + 
  geom_bar(aes(x=reorder(hotel_cluster,num),y=num,fill=user_location_country),stat = "identity") + 
  scale_fill_brewer(palette = "RdYlBu") +
  facet_wrap(~user_location_country,scales= "free") +
  
  xlab("Hotel Cluster") + ylab("Frequency") + 
  labs(fill="Country ID")+
  ggtitle("Top 5 Popular Hotel Clusters",
  subtitle = "For Users in Top 5 Countris for Number of Hotel Booking per User,\nFrom Jan. 2013 to Dec. 2014") + 
  coord_flip()
```

#### Which destination the users like most?

```{r message=FALSE, warning=FALSE}
df %>% 
  dplyr::select(is_booking,srch_destination_id) %>% 
  group_by(srch_destination_id) %>% 
  summarise(num=n()) %>% arrange(desc(num)) %>% 
  
  head(10) %>% ggplot() + 
  geom_point(aes(x=reorder(srch_destination_id,num),y=num,size=num),alpha=0.7,col="royalblue") + 
  scale_size(range = c(3,9.5)) +
  xlab("Destination ID") + ylab("Frequency") + 
  ggtitle("Top 10 Popular Destinations in Hotel Search",
          subtitle = "From Jan. 2013 to Dec. 2014") +
  guides(size=F)
```

### 2.1.3 Expedia-related

#### Daily Active User
```{r message=FALSE, warning=FALSE}
# Daily Active User
df %>% mutate(date=ymd(as.Date(df$date_time))) %>%
  dplyr::select(date,posa_continent,is_booking) %>% 
  group_by(date,posa_continent) %>% 
  summarise(user_num=n()) %>% ungroup() %>% 
  # Draw line chart
  ggplot() + 
  geom_line(aes(x=date,y=user_num,col=posa_continent)) +
  scale_color_brewer(palette = "Set2")+
  xlab("Date") + ylab("Number of User") +
  labs(col="Continent ID \nof the Website")+
  ggtitle("Daily Active User on Expedia by Continent",subtitle = "From Jan. 2013 to Dec. 2014")
```

#### Booking Channel

```{r message=FALSE, warning=FALSE}
df %>% dplyr::filter(is_booking==1) %>% 
  dplyr::select(channel,is_booking) %>% 
  group_by(channel) %>% 
  summarise(num=n()) %>% 
  top_n(5,num) %>% 
  arrange(desc(num)) %>% 
  
  ggplot()+
  geom_bar(aes(x=reorder(channel,num),y=num,fill=channel),stat = "identity")+
  scale_fill_brewer(palette = "Set2")+
  xlab("Channel ID") + ylab("Number of Booking")+
  ggtitle("Top 5 Channel of Hotel Booking",subtitle = "From Jan. 2013 to Dec. 2014")+
  labs(fill="Channel ID")+
  coord_flip()
```

#### How many user used package to book hotel?
```{r message=FALSE, warning=FALSE}
df %>% dplyr::filter(is_booking==1) %>% 
  dplyr::select(is_package,hotel_cluster) %>% 
  filter(hotel_cluster %in% c("91","48","42","59","28")) %>% 
  group_by(hotel_cluster,is_package) %>% 
  summarise(num=n()) %>% 
  
ggplot()+
  geom_bar(aes(x=as.character(is_package),y=num,fill=hotel_cluster),stat="identity")+
  facet_wrap(~hotel_cluster,scales = "free")+
  scale_fill_brewer(palette = "Set2") +
  xlab("Package Using Status")+ylab("Frequency")+
  labs(fill="Hotel Cluster") +
  ggtitle("Package Using Status of Top 5 Popular Hotel Clusters",subtitle = "From Jan. 2013 to Dec. 2014") + 
  coord_flip()
```

#### How many user used mobile to book hotel?
```{r message=FALSE, warning=FALSE}
df %>% dplyr::filter(is_booking==1) %>% 
  dplyr::select(is_mobile,hotel_cluster) %>% 
  filter(hotel_cluster %in% c("91","48","42","59","28")) %>% 
  group_by(hotel_cluster,is_mobile) %>% 
  summarise(num=n()) %>% 
  
ggplot()+
  geom_bar(aes(x=as.character(is_mobile),y=num,fill=hotel_cluster),stat="identity")+
  facet_wrap(~hotel_cluster,scales = "free")+
  scale_fill_brewer(palette = "Set2") +
  xlab("Mobile Using Status")+ylab("Frequency")+
  labs(fill="Hotel Cluster") +
  ggtitle("Mobile Using Status of Top 5 Popular Hotel Clusters",subtitle = "From Jan. 2013 to Dec. 2014") + 
  coord_flip()
```

## 2.2 Feature Engineering

### 2.2.1 Feature Establishment
 - Duration of Stay: stay_duration
Since `srch_ci` denotes check-in date and `srch_co` denotes check-out date, the duration of stay would be a feature of interest to explore users' hotel booking behavior.

```{r}
# Stay Duration
df2 <- df %>% 
  mutate(stay_duration=
           as.Date(df$srch_co)-as.Date(df$srch_ci)) %>% 
  dplyr::select(-c(srch_ci,srch_co)) %>%
  mutate(stay_duration=as.numeric(stay_duration))
```

 - Hotel Rank: hotel_rank
Meanwhile, since the number of hotel cluster is relatively large (100 hotel clusters), we would like to transform it into a 10-point scale based on the number of booking for the simplicity of the analysis. 

If hotel cluster ID equals to top 10 frequently booked hotel, then it would be ranked as 10. If hotel cluster ID equals to top 11-20 frequently booked hotel, it would be ranked as 9.If hotel cluster ID equals to top 11-30 frequently booked hotel, it would be ranked as 8. The rest of the ranking will do as above. 
```{r message=FALSE, warning=FALSE}
hotel_booking <- df %>% 
  dplyr::select(is_booking,hotel_cluster) %>% 
  dplyr::filter(is_booking == 1) %>% 
  group_by(hotel_cluster) %>% 
  summarise(num=n()) %>% 
  arrange(desc(num))
```


```{r}
# Hotel ranking
df2$hotel_rank <- rep(NA,nrow(df2))

df2$hotel_rank <- ifelse(df2$hotel_cluster%in% hotel_booking$hotel_cluster[1:10],10,
                         ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[11:20],9,
                                ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[21:30],8,
                                       ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[31:40],7,
                                              ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[41:50],6,
                                                     ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[51:60],5,
                                                            ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[61:70],4,
                                                              ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[71:80],3,
                                                              ifelse(df2$hotel_cluster%in%hotel_booking$hotel_cluster[81:90],2,1)))))))))

```


### 2.2.2 Feature Selection

Since **site_name** and **posa_continent** in the dataset all reflect which continent the users are based, and posa_continent has similar structure as **hotel_continent**, the posa_continent will be chosen for the modelling part.

Meanwhile, if the user has children that are needed to be placed in a seprate room, it will reflect in the **srch_rm_cnt**, so we will choose this column and drop **srch_adults_cnt** and **srch_children_cnt**.

Because the meaning of **hotel_market** and **srch_destination_type_id** are not specified in the data description part on Kaggle, those columns will also be dropped for interpretability of the variable.

What's more, user has city-level feature **user_location_city** and region-level feature **user_location_region**, however, hotel only have continent feature **hotel_continent** and country-level feature **hotel_country**, so **user_location_city** and **user_location_region** will be dropped for the correpondence of the features. Morover,**date_time** will also be dropped for the simplicity of the model.


```{r}
df_final <- df2 %>%
  dplyr::filter(is_booking==1) %>% 
  dplyr::select(-c(hotel_cluster,date_time,is_booking,hotel_market,site_name,srch_children_cnt,srch_adults_cnt,user_location_city,user_location_region,srch_destination_type_id)) 
```

```{r}
set.seed(678)
df_final <- df_final %>% sample_frac(0.5,replace = F)
```

```{r}
colnames(df_final)
```

# 3 Model
## 3.1 Model Establishment
```{r}
# fit the model
fit <- polr(ordered(hotel_rank)~user_location_country + posa_continent + orig_destination_distance  + hotel_continent+ is_mobile + is_package  + srch_rm_cnt  + cnt + channel+  stay_duration,data=df_final,Hess = T)
```

```{r}
display(fit)
```

Visualize the model
```{r}
predx<-expand.grid(
  user_location_country=unique(df_final$user_location_country),
  posa_continent=unique(df_final$posa_continent),
  orig_destination_distance=mean(df_final$orig_destination_distance),
  hotel_continent=unique(df_final$hotel_continent),
  is_mobile=0,is_package=0,
  srch_rm_cnt=mean(df_final$srch_rm_cnt),
  cnt=mean(df_final$cnt),
  channel=unique(df_final$channel),
  stay_duration=mean(df_final$stay_duration))
predy<-predict(fit,newdata=predx,type="prob")
```

```{r message=FALSE, warning=FALSE}
resd<-data.frame(predx[,c("user_location_country","posa_continent","orig_destination_distance","hotel_continent","is_mobile","is_package","srch_rm_cnt","cnt","channel","stay_duration")],hotel_rank=predy)

ggplot(melt(resd,id.var=c(c("user_location_country","posa_continent","orig_destination_distance","hotel_continent","is_mobile","is_package","srch_rm_cnt","cnt","channel","stay_duration"))))+
geom_bar(position = "fill",stat="identity")+
aes(x=hotel_continent,y=value,fill=variable)+
facet_grid(~posa_continent)+
theme(axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(angle = 45, hjust = 1))+
ggtitle("Probability by posa_continent")
```

## 3.2 Model Checking
```{r}
obsmat <- model.matrix(~ordered(hotel_rank)-1,data=df_final)
resdimat<-obsmat-fitted(fit)
```

```{r}
par(mfrow=c(2,5))
binnedplot(fitted(fit)[,1],resdimat[,1])
binnedplot(fitted(fit)[,2],resdimat[,2])
binnedplot(fitted(fit)[,3],resdimat[,3])
binnedplot(fitted(fit)[,4],resdimat[,4])
binnedplot(fitted(fit)[,5],resdimat[,5])
binnedplot(fitted(fit)[,6],resdimat[,6])
binnedplot(fitted(fit)[,7],resdimat[,7])
binnedplot(fitted(fit)[,8],resdimat[,8])
binnedplot(fitted(fit)[,9],resdimat[,9])
binnedplot(fitted(fit)[,10],resdimat[,10])
```

